Continuous Deductive Reasoning over Streamed RDF Data

### 2015-06-26 content/contribution/R4.tex
fixed the graphics inclusion in R4 to explicitly include 'png's
added the new text from Rehab, and added references to the operators of CoDeR and semantics of Continuous Datalog, where appropriate.  Also modified the EBNF to reflect the windowing as actually implemented, based on the description of the implementation.

\begin{nestedsection}{R4: Rule-based Reasoning over RDF streams using Rete}{implementation}
	As a proof of concept, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
	Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams.
	The system architecture is shown in \reffig{R4-architecture}.
	\begin{figure}[b]
		\centering
		\includegraphics[width=0.45\textwidth]{R4-architecture}
		\caption{The system architecture of R4.}
		\labelfig{R4-architecture}
	\end{figure}

	A rule document containing any number of rules is submitted to the system.
	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards.
	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in \reffig{lst-S-RIF-Core}.

	\begin{figure*}[t]
		\centering
		\begin{ebnf}
			\Rule{Document}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Document} 
			        \Token{(} 
			        \Optional{\NTerm{Base}} 
			        \ZeroOrMore{\NTerm{Prefix}} 
			        \ZeroOrMore{\NTerm{Import}} 
			        \Optional{\NTerm{Group}} 
			        \Token{)}
			}
			\Rule{Base}{
			        \Token{Base} 
			        \Token{(} 
			        \NTerm{ANGLEBRACKIRI} 
			        \Token{)}
			}
			\Rule{Prefix}{
			        \Token{Prefix} 
			        \Token{(} 
			        \NTerm{Name} 
			        \NTerm{ANGLEBRACKIRI} \Token{)}
			}
			\Rule{Import}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Import} 
			        \Token{(} 
			        \NTerm{LOCATOR} 
			        \Optional{\NTerm{PROFILE}} 
			        \Token{)}
			}
			\Rule{Group}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Group} 
			        \Token{(} 
			        \ZeroOrMore{\OrList{\NTerm{RULE} \OrSep \NTerm{Group}}} 
			        \Token{)}
			}
			\Rule{RULE}{
			        \Optional{\NTerm{IRIMETA}}
			        \Token{Forall}
			        \OneOrMore{\NTerm{Variable}}
			        \Token{(}
			        \NTerm{CLAUSE}
			        \Token{)}
			}
			\OrRule{
			        \NTerm{CLAUSE}
			}
			\Rule{CLAUSE}{
			        \OrList{\NTerm{Implies} \OrSep \NTerm{ATOMIC}}
			}
			\Rule{Implies}{
			        \Optional{\NTerm{IRIMETA}}
			        \OrList{\NTerm{ATOMIC} \OrSep \Token{And} \Token{(} \ZeroOrMore{ATOMIC} \Token{)}}
			        \Token{:-}
			        \NTerm{FORMULA}
			}
			\Rule{LOCATOR}{
			        \NTerm{ANGLEBRACKURI}
			}
			\Rule{PROFILE}{
			        \NTerm{ANGLEBRACKURI} 
			        \Optional{\NTerm{Window}}
			}
			\Rule{Window}{
			        \NTerm{Number}
			        \NTerm{TimeUnit}
			}
			\Rule{TimeUnit}{
			        \OrList{\Token{ms} \OrSep \Token{s} \OrSep \Token{m} \OrSep \Token{h}}
			}
		\end{ebnf}
		\caption{Streaming RIF-Core Grammar}
		\labelfig{lst-S-RIF-Core}
	\end{figure*}

@@

		However, the Rete algorithm is memory-intensive as it stores all partial matches, trading memory for speed.
		In a streaming context, this is infeasible as streams can grow without bounds.
		Therefore, we place time constraints on the memories using a window join operator instead of the traditional joins.

		Talk about operators and link them to CoDeR.

@@

 		\begin{figure*}[t]
 			\centering
 			\includegraphics[width=0.9\textwidth]{example-rete-network}
 			\caption{An example Rete network with corresponding (RDFS) rules.}
 			\labelfig{example-rete-network}
 		\end{figure*}
	\end{nestedsection}
\end{nestedsection}

>>>>

\begin{nestedsection}{R4: Rule-based Reasoning over RDF streams using Rete}{implementation}
	As a proof of concept, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
	Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams.
	The system architecture is shown in \reffig{R4-architecture}.
	\begin{figure}[b]
		\centering
		\includegraphics[width=0.45\textwidth]{R4-architecture.png}
		\caption{The system architecture of R4.}
		\labelfig{R4-architecture}
	\end{figure}

	A rule document containing any number of rules is submitted to the system.
	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards.
	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in \reffig{lst-S-RIF-Core}.
	Further, we propose the extension of the set of PROFILEs that describe the IMPORTs of a rule set to include those definitions of RDF streams recognised by the W3C.

	\begin{figure*}[t]
		\centering
		\begin{ebnf}
			\Rule{Document}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Document} 
			        \Token{(} 
			        \Optional{\NTerm{Base}} 
			        \ZeroOrMore{\NTerm{Prefix}} 
			        \ZeroOrMore{\NTerm{Import}}
			        \Optional{\NTerm{Window}}
			        \Optional{\NTerm{Group}} 
			        \Token{)}
			}
			\Rule{Base}{
			        \Token{Base} 
			        \Token{(} 
			        \NTerm{ANGLEBRACKIRI} 
			        \Token{)}
			}
			\Rule{Prefix}{
			        \Token{Prefix} 
			        \Token{(} 
			        \NTerm{Name} 
			        \NTerm{ANGLEBRACKIRI} \Token{)}
			}
			\Rule{Import}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Import} 
			        \Token{(} 
			        \NTerm{LOCATOR} 
			        \Optional{\NTerm{PROFILE}} 
			        \Token{)}
			}
			\Rule{Group}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Group} 
			        \Token{(} 
			        \ZeroOrMore{\OrList{\NTerm{RULE} \OrSep \NTerm{Group}}} 
			        \Token{)}
			}
			\Rule{RULE}{
			        \Optional{\NTerm{IRIMETA}}
			        \Token{Forall}
			        \OneOrMore{\NTerm{Variable}}
			        \Token{(}
			        \NTerm{CLAUSE}
			        \Token{)}
			}
			\OrRule{
			        \NTerm{CLAUSE}
			}
			\Rule{CLAUSE}{
			        \OrList{\NTerm{Implies} \OrSep \NTerm{ATOMIC}}
			}
			\Rule{Implies}{
			        \Optional{\NTerm{IRIMETA}}
			        \OrList{\NTerm{ATOMIC} \OrSep \Token{And} \Token{(} \ZeroOrMore{ATOMIC} \Token{)}}
			        \Token{:-}
			        \NTerm{FORMULA}
			}
			\Rule{LOCATOR}{
			        \NTerm{ANGLEBRACKURI}
			}
			\Rule{PROFILE}{
			        \NTerm{ANGLEBRACKURI} 
			}
			\Rule{Window}{
			        \NTerm{Number}
			        \NTerm{TimeUnit}
			}
			\Rule{TimeUnit}{
			        \OrList{\Token{ms} \OrSep \Token{s} \OrSep \Token{m} \OrSep \Token{h}}
			}
		\end{ebnf}
		\caption{Streaming RIF-Core Grammar}
		\labelfig{lst-S-RIF-Core}
	\end{figure*}

@@

		However, the Rete algorithm is memory-intensive as it stores all partial matches, trading memory for speed.
		In a streaming context, this is infeasible as streams can grow without bounds.
		Therefore, we place time constraints on the memories using a window-join operator instead of the traditional joins.

		Data enters the network through the source nodes, which convert streams of graphs into streams of quads, by extracting the graph ID and attaching it to every triple in this graph.
		Source nodes are, therefore, implementations of the graph-to-stream operator ${{}_g{\pi_t}}$ of CoDeR.
		Source nodes also time stamps the incoming tuples with system timestamp if they are not already stamped from their origins.
		They propagate the data to the successor nodes, which are the alpha nodes.

		Alpha nodes are single-entry nodes that form a discrimination network.
		In R4, alpha nodes can receive input from any number of nodes through its single entry point.
		Each input triple is matched against some conditions, and is either dropped if there is no match, or propagated downstream to the successor nodes in the case of a match.
		As such, R4's interpretation of alpha nodes are implementations of the triple-select operator $\sigma_{triple}$ of CoDeR, but can also filter based on the datatypes of the values in triples, thereby also implementing the other selection operator of CoDeR, $\sigma_{external}$.
		The first alpha node in the network propagates its output to a special node called the left input adaptor, while the other alpha nodes propagate their output to join nodes.
		The left input adaptor node is responsible of preparing the incoming triples to enter the beta network from the left input of the first join node in a given join-tree.
		It creates a new token i.e. a partial match, defined as a list of triples, for each incoming triple and adds this triple as the first item in the list.
		Tokens are then sent to the first join node in the beta network.

		Beta nodes are two-entry nodes that are responsible for joining the branches of the alpha network.
		In our implementation, as in the original Rete algorithm, beta nodes form a left-deep tree.
		Each beta node maintains a left memory, which is a beta memory storing tokens received from other beta nodes (or from the left input adaptor node in case of the first beta node) and a right memory, which is an alpha memory storing triples received from alpha nodes.
		Join nodes are beta nodes that match tuples from both sides according to some conditions, e.g. a shared variable binding.
		As explained earlier, we use stream-to-stream window-join operators to avoid storing and operating over all partial results.
		In this context, each left (beta) and right (alpha) memory is implemented as a \emph{valid window}.
		These windows are implemented as priorities queues and their items are ordered according to their timestamps.

		When a join node is left activated, i.e. it receives a token through its left input, it first adds the new token to its left window, then it iterates over its right window elements trying to find a match.
		When a match is found, a new token is created by adding the right triple to the triples list of the left token.
		The new token is stamped with the earliest timestamp of its triples, and then propagated to the next join node or to the terminal node if it is the last join.
		At this point, the join node performs a garbage collection over its left window, ensuring that items with timestamps that are smaller that current time minus the semantic window size are deleted.
		Conversely, when the join node is right activated by receiving a triple from an alpha node, it adds the new triple to the right window, matches it against the elements of the left window, and performs a garbage collection over the right window.

		This method of annotating tokens with the earliest timestamp of its components initially appears to violate the annotation semantics laid out in \refsec{semantics}.
		However, the ordering of tokens in their streams and their propagation through the system expresses those annotation semantics implicitly.
		This leaves the timestamp of a token, in combination with the size of the semantic window associated with the rule set, able to express the negation time $i_n$ of instances in CoDeR.
		As such, the interpretation of beta nodes in R4 is an implementation of the window-join operator characterised as a pair of SteMs implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.

 		Finally, terminal nodes receive tokens from the last node in the network and are responsible of producing the new entailed triples.

@@

 		\begin{figure*}[t]
 			\centering
 			\includegraphics[width=0.9\textwidth]{example-rete-network.png}
 			\caption{An example Rete network with corresponding (RDFS) rules.}
 			\labelfig{example-rete-network}
 		\end{figure*}
	\end{nestedsection}
\end{nestedsection}
